How predictable is extinction? Forecasting species survival at million-year timescales

### 2019-04-15 doc/paper.tex
work smithing before sending to seth

We find that all of our models are expected to correctly forecasting the rank order in extinction risk of two randomly selected future observations between 70\% to 80\% of the time. One of the most striking aspects of these results is the similarity in forecasting performance for in-sample and out-of-sample observations. A slight decrease in performance when dealing with out-of-sample observations makes sense: each of the models fit during cross-validation is based on fewer data than the model fit on the full data (between 1/5th to 4/5ths of the original). Additionally, a potential decrease in precision when forecasting the future of extinction risk is almost to be expected as the future is not necessarily the same as the past. However, the similarly between the in-sample and out-of-sample results indicates that our model is fairly robust to how extinction intensity has changed over the Cenozoic. An important caveat to our results is that human impacts may substantially alter extinction risk dynamics away from the average Cenozoic ``state'' so that the future could become less predictable than it has been in the past.

>>>>

We find that all of our models are expected to correctly forecasting the rank order in extinction probabilities for a randomly selected extinct-extant pair of future observations between 70\% to 80\% of the time (Fig. \ref{fig:fold_auc}). One of the most striking aspects of these results is the similarity in forecasting performance for in-sample and out-of-sample observations. A slight decrease in performance when dealing with out-of-sample observations makes sense: each of the models fit during cross-validation is based on fewer data than the model fit on the full data (between 1/5th to 4/5ths of the original). Additionally, a potential decrease in precision when forecasting the future of extinction risk is almost to be expected as the future is not necessarily the same as the past. However, the similarly between the in-sample and out-of-sample results indicates that our model is fairly robust to how extinction intensity has changed over the Cenozoic. An important caveat to our results is that human impacts may substantially alter extinction risk dynamics away from the average Cenozoic ``state'' so that the future could become less predictable than it has been in the past.
